Analysis of Bauer et al. (unpublished) Field experiment: <br> Persistence

### 2022-12-05 R/model_persistence_check.Rmd
Update
Create model_persistence_check.Rmd

---
title: "Persistence of seed mixture species: <br> Analysis of Bauer et al. (unpublished) Field experiment"
author: "Markus Bauer <br>"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Persistence sensu Wilsey (2021) https://doi.org/10.1111/rec.13132

@@

```{r message = FALSE, tidy = TRUE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
library(DHARMa)
library(DHARMa.helpers)
library(bayesplot)
library(loo)
library(tidybayes)
library(emmeans)

@@

```{r}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
      .default = "?",
      plot = "f",
      site = "f",
      sand_ratio = "f",
      substrate_depth = col_factor(levels = c("30", "15")),
      target_type = col_factor(levels = c(
        "hay_meadow", "dry_grassland"
      )),
      seed_density = "f",
      exposition = col_factor(levels = c(
        "north", "south"
      )),
      survey_year = "c"
    )
  ) %>%
  ### Exclude data of seed mixtures
  filter(presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    id = factor(id),
    n = persistence
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, n
  )
```

@@

```{r}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

## Model building

### Models

Possible prior distributions

```{r}
ggplot(data = data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 10)) +
  expand_limits(y = 0) + ggtitle("Normal distribution")
ggplot(data = data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 10)) +
  expand_limits(y = 0) + ggtitle("Cauchy distribution")
ggplot(data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 10)) +
  expand_limits(y = 0) + ggtitle(expression(Student~italic(t)*"-distribution"))
```

Specifications for the models

```{r}
iter = 10000
chains = 4
thin = 2
priors <- c(
  set_prior("normal(0, 20)", class = "b"),
  set_prior("normal(-2.5, 20)", class = "b", coef = "sand_ratio25"),
  set_prior("normal(-5, 20)", class = "b", coef = "sand_ratio50"),
  set_prior("normal(5, 20)", class = "b", coef = "expositionsouth"),
  set_prior("normal(-2.5, 20)", class = "b", coef = "survey_year_fct2019"),
  set_prior("normal(-5, 20)", class = "b", coef = "survey_year_fct2020"),
  set_prior("normal(-7.5, 20)", class = "b", coef = "survey_year_fct2021"),
  set_prior("cauchy(0, 10)", class = "sigma")
)
```

Model caluclations

```{r}
load(file = here("outputs", "models", "model_persistence_1.Rdata"))
load(file = here("outputs", "models", "model_persistence_3.Rdata"))
```

### Model comparison

```{r}
m_1 <- m1
m_2 <- m3
m_1$formula
m_2$formula
```

Conditional R² values

```{r}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot))
```

Marginal R² values

```{r}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

### Model check

#### DHARMa

```{r}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

#### Preparation

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

#### Samling efficency/effectiveness (Rhat and EFF)

```{r}
range(draws1$rhat)
range(draws2$rhat)
range(draws1$ess_bulk)
range(draws2$ess_bulk)
range(draws1$ess_tail)
range(draws2$ess_tail)
```

#### MCMC diagnostics

@@

Priors and conditional and marignal R²

```{r}
prior_summary(m_1, all = FALSE)
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

Posteriors

```{r}
draws1
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
```

@@

```{r}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
```

>>>>

---
title: "Analysis of Bauer et al. (unpublished) Field experiment: <br> Persistence"
author: "<b>Markus Bauer*</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
    df_print: tibble
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

Technichal University of Munich, TUM School of Life Sciences, Chair of Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

<span>&#42;</span> Corresponding author: [markus1.bauer@tum.de](mailto:markus1.bauer@tum.de)
<br>
ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)

Persistence sensu Wilsey (2021) Restor Ecol [DOI: 10.1111/rec.13132](https://doi.org/10.1111/rec.13132)

@@

```{r message = FALSE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
library(DHARMa)
library(DHARMa.helpers)
library(bayesplot)
library(loo)
library(tidybayes)
library(emmeans)
```

@@

```{r}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""),
  col_types = cols(
    .default = "?",
    plot = "f",
    site = "f",
    sand_ratio = "f",
    substrate_depth = col_factor(levels = c("30", "15")),
    target_type = col_factor(levels = c("hay_meadow", "dry_grassland")),
    seed_density = "f",
    exposition = col_factor(levels = c("north", "south")),
    survey_year = "c"
  )
  ) %>%
  ### Exclude data of seed mixtures
  filter(presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, exposition, sep = " "),
    botanist_year = factor(botanist_year),
    id = factor(id),
    n = persistence
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
  )
```

@@

```{r echo = FALSE}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

## Models

```{r collapse = TRUE}
load(file = here("outputs", "models", "model_persistence_1.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
m_1 <- m1
m_2 <- m_full
```

```{r collapse = TRUE}
m_1$formula
m_2$formula
```

```{r collapse = TRUE}
m_1$family
m_2$family
```

### Priors

#### Possible prior distributions

```{r}
get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site/plot),
          data = sites)
ggplot(data = data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 10)) +
  expand_limits(y = 0) + ggtitle("Normal distribution")
ggplot(data = data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 10)) +
  expand_limits(y = 0) + ggtitle("Cauchy distribution")
ggplot(data.frame(x = c(-40, 40)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 10)) +
  expand_limits(y = 0) + ggtitle(expression(Student~italic(t)*"-distribution"))
```

#### Prior summary

```{r}
prior_summary(m_1, all = FALSE)
```

Conditional <i>R</i>² values

```{r collapse = TRUE}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year))
```

Marginal <i>R</i>² values

```{r collapse = TRUE}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

## Model check

### DHARMa

```{r collapse = TRUE}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

### Preparation

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

### Sampling efficency and effectiveness

#### Rhat

```{r collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

#### Effective sampling size (ESS)

```{r collapse = TRUE}
mcmc_neff(neff_ratio(m_1))
mcmc_neff(neff_ratio(m_2))
```

### MCMC diagnostics

@@

Conditional and marignal <i>R</i>²

```{r collapse = TRUE}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

Posteriors of chosen model

```{r}
draws1
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

Posteriors of second model:

```{r}
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

@@

Just, to get exact values if necessary, which is not possible from the figure

```{r max_height = "1000px"}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
```

```{r echo = FALSE}
write.csv(draws1, here("outputs", "statistics", "table_fcs.csv"))
```

# Session info

```{r echo = FALSE, max_height = "500px"}
sessionInfo()
```
